Modeling site-specific amino-acid preferences deepens phylogenetic estimates of the divergence of viral proteins.

### 2018-03-13 manuscript/manuscript.tex
small wording tweaks

\begin{figure}
\centerline{\includegraphics[width=0.99\textwidth]{figures/phylobayes.pdf}}
\caption{\label{fig:phylobayes}
\textbf{Models inferred from natural sequences have similar stationary states to models defined by experimental preferences and estimate similar branch lengths.}
(A) We fit an ExpCM(H1+H3 avg)+$\Gamma\omega$ and a pbMutSel to the full HA tree in \ref{fig:empirical_trees}. 
The pbMutSel amino-acid preferences are inferred from the natural HA sequences, while the ExpCM amino-acid preferences are experimentally measured and then rescaled by the stringency parameter in \ref{tab:empirical_data}. 
The pbMutSel preferences are more correlated with the re-scaled average of the H1 and H3 deep mutational scanning preferences than the individual re-scaled H1 and H3 deep mutational scanning preferences are to each other (Pearson's $r$: 0.74 versus 0.52). 
(B) The ExpCM(H1+H3 avg)+$\Gamma\omega$ and pbMutSel models estimated similar branch lengths when fit to the entire HA tree. 
\jdbcomment{You need to mention and give a bit of explanation for the coloring. What does it mean for a branch to be ``from'' H1 or H3? Would it be better to say branch ``leading to''? And does that mean leading to it with respect to the root of the tree as shown in \ref{fig:empirical_trees}?}
The \texttt{phydms} program implementing ExpCM's and the \texttt{phylobayes} program implementing pbMutSel models give branch lengths in different units, so to facilitate direct comparison between the models, we have normalized all branch lengths returned by each program by the length of the branches separating the earliest (\jdbcomment{say exact strain}) and latest (\jdbcomment{say exact strain}) seasonal human H1 sequences on the three. 
}
\end{figure}

The previous sections used ExpCM's, which are mutation-selection models that use site-specific amino-acid preferences that have been measured by experiments. 
However, there are other mathematically similar implementations of mutation-selection models that infer the amino-acid preferences directly from the natural sequence data. 
When these models are designed for use in phylogenetic inference, they are generally implemented in a Bayesian framework, which avoids the overfitting problems associated with trying to make maximum-likelihood estimates of the thousands of amino-acid preference parameters~\citep{lartillot2014overcoming}.
(Note that the maximum-likelihood implementations of \citet{tamuri2012estimating,tamuri2014penalized} are designed for estimating the amino-acid preferences, \emph{not} for phylogenetic inference.)
The model most comparable to our ExpCM's is the codon mutation-selection model implemented in \texttt{phylobayes}, which we will refer to as pbMutSel~\citep{rodrigue2014site}. 
\jdbcomment{Throughout this section, you refer to \texttt{phylobayes}. But are you sure that \texttt{phylobayes} is the same thing as the \texttt{PhyloBayes-MPI} program that implements the pbMutSel models? I didn't look into this in great detail, but I got different websites when I Googled the two. Check this, and if it's actually \texttt{PhyloBayes-MPI} then update all references to the program to indicate that.}
In the pbMutSel model, the amino-acid preferences are modeled using Dirichlet processes rather than derived from experiments. 
However, like an ExpCM, a pbMutSel model still assumes a single set of time-homogeneous site-specific amino-acid preferences for the entire tree.

@@

We fit a pbMutSel model to the entire HA phylogenetic tree, and compared the results to those from analyzing the same tree with the best ExpCM, which is the ExpCM(H1+H3 avg)+$\Gamma\omega$ variant.
This is a direct apple-to-apples comparison, since the pbMutSel model also draws $\omega$ from a gamma-distribution.
\jdbcomment{Is the previous sentence correct? If so, can you add citation for it?}
First, we compared the amino-acid preferences inferred by the pbMutSel model to the preferences measured in the experiments.
\ref{fig:phylobayes}A shows that the preferences inferred by pbMutSel are quite similar to the (H1+ H3 avg) obtained by averaging the deep mutational scanning measurements for the H1 and H3 HAs. 
Notably, the amino-acid preferences from the pbMutSel model are more correlated with the (H1+ H3 avg) than the H1 and H3 measurements are with each other (\ref{fig:phylobayes}A).
This strong correlation indicates that the ExpCM(H1+H3 avg)+$\Gamma\omega$ is unlikely to be much different than a pbMutSel model that is parameterized only using the natural sequence data.

@@

We created four subalignments with ``low" and ``intermediate" divergence from either the H1 or the H3 deep mutational scanning reference sequence \jdbcomment{for the analysis in Figure ?}. 
The ``low divergence" alignments had $\ge 59\%$ amino-acid identity to the sequence used in the deep mutational scanning, and the ``intermediate divergence" alignments had $\ge 46\%$ identity from the reference sequence.
The resulting alignments are found in \ref{suppfile:alignment_lowH1}, \ref{suppfile:alignment_lowH3}, \ref{suppfile:alignment_intermediateH1}, and \ref{suppfile:alignment_intermediateH3}.  
\jdbcomment{Again, we should probably get rid of these four preceding files since we have too many. Presumably they can be re-generated using the description above of the identity cutoffs?}

@@

\subsection*{pbMutSel inference with \texttt{phylobayes}.}
For \ref{fig:phylobayes}, we fit a pbMutSel model to the full HA tree. 
We ran one chain for 5000 steps, saved every sample, and discarded the first 500 samples as a burnin. 
\jdbcomment{Presumably the program gives some measure of convergence? If so, describe what it was.}
\jdbcomment{You need to better describe what the tree represents and what the pbMutSel preferences represent. Are they the posterior mean?}

In order to make the branch lengths in \ref{fig:phylobayes} comparable between the pbMutSel tree returned by \texttt{phylobayes} and the other trees returned by \texttt{phyddms}, we we normalized the branch lengths on the pbMutSel consensus tree and the ExpCM(H1+H3 avg)+$\Gamma\omega$ by dividing each branch by the length from A/South Carolina/1/1918 and A/Solomon Islands/3/2006. 
These two H1 sequences are early and late representatives of the longest known human influenza lineage, and are of sufficiently high identity that different ExpCM and GY94 substitution models all estimate nearly identical branch lengths separating them.

\subsection*{Software versions and computer code}
All code used for the analyses in this paper is available at \url{https://github.com/jbloomlab/divergence_timing_manuscript}. 
The external computer programs that we used were
\begin{itemize} 
\item \texttt{phydms}~\citep{hilton2017phydms} version 2.2.2 (available at \url{github.com/jbloomlab/phydms}) to fit the ExpCM and GY94 models.
\item \texttt{pyvolve}~\citep{spielman2015pyvolve} version 0.8.7 (available at \url{https://github.com/sjspielman/pyvolve}) to simulate the sequences.
\item \texttt{phylobayes}~\citep{rodrigue2014site} version 1.8 (available at \url{https://github.com/bayesiancook/pbmpi}) to fit the pbMutSel model. 
\item \jdbcomment{RAxML}
\item We used \texttt{ggplot2}~\citep{wickham2016ggplot2}, \texttt{ggtree}~\citep{yu2017ggtree}, and \texttt{ggseqlogo}~\citep{wagih2017ggseqlogo} for visualization of the results.
\end{itemize}

@@

I would say keep the map of site numbers, the H1 and H3 preferences, and the alignment. 
I think all the other files can either be re-generated using the other ones if you provide adequate instructions in the methods (e.g., averaging the preferences or pruning the full alignment to sub-alignments), or simply have model parameters which could be put in a supplementary table.

Also, put the supplementary information after the references.

Also, put a page break after each supplementary figure / table using the \texttt{clearpage} command.
}

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_doud}}
\caption{\label{suppfig:prefs_doud}
\textbf{H1 HA amino-acid preferences measured by deep mutational scanning.}
Each column represents a site in the HA protein, and the height of each letter is proportional to the preference for the amino acid measured by~\citet{doud2016accurate} and then re-scaled by the stringency parameter in \ref{tab:empirical_data}. 
The plot only shows sites that are alignable between the H1 and H3 HAs, and these alignable sites are numbered sequentially starting from 1.
The conversion between the numbering scheme in this figure and sequential numbering of the H1 HA reference sequence is in \ref{suppfile:WSN_Perth_map}. 
\jdbcomment{For this figure and the other logo plots, increase the number of rows by a few so things are a bit bigger.}
}
\end{suppfig}

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_lee}}
\caption{\label{suppfig:prefs_lee}
\textbf{H3 HA amino-acid preferences measured by deep mutational scanning.}
Similar to \ref{suppfig:prefs_doud} but shows the re-scaled preferences for the H3 HA as measured by \citet{lee2018deep}.
 }
\end{suppfig}

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_average}}
\caption{\label{suppfig:prefs_average}
\textbf{Average of H1 HA  and H3 HA amino-acid preferences measured by deep mutational scanning.}
Similar to \ref{suppfig:prefs_doud} but shows the re-scaled average of the preferences for the H1 and H3 HAs.
}
\end{suppfig}

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_mutSel}}
\caption{\label{suppfig:prefs_mutSel}
\textbf{Amino-acid preferences inferred by the pbMutSel model.}
Similar to \ref{suppfig:prefs_doud}, but shows the preferences inferred by fitting the pbMutSel model to the full HA tree.
}
\end{suppfig}

\begin{suppfig}[H]
\centerline{\includegraphics[width=0.50\textwidth]{figures/divergence_distances.pdf}}
\caption{\label{suppfig:subalignments}
\textbf{Overall divergence for the subtrees in \ref{fig:compete}.} 
\jdbcomment{This legend is indecipherable because rather than saying what you are showing you just refer to a lot of other figures / files.
Reduce the references to other figures / files to the minimum needed an explain what is actually being shown.}
We created two sub alignments from the alignment (\ref{suppfile:alignment_high}) used in \ref{fig:empirical_trees} for each deep mutational scanning reference HA. 
The ``low" alignments have $\ge59\%$ amino-acid identity to the reference sequence (\ref{suppfile:alignment_lowH1}, \ref{suppfile:alignment_lowH3}) and the ``intermediate" alignments have $\ge46\%$ amino-acid identity to the reference sequence (\ref{suppfile:alignment_intermediateH1}, \ref{suppfile:alignment_intermediateH3}).
}
\end{suppfig}

\begin{supptable}[t!]
\caption{\label{supptab:tree_lengths}
{\bf Total branch length for the trees in \ref{fig:empirical_trees}.}
The branch lengths are in average codon substitutions per site. 
\jdbcomment{Change to be tree diameter, which is distance between two most distant tip nodes.
Also, to make it easier to relate back to the percentage increases in the text, make 8 rows, one for each model and get rid of the ``rate variation'' column. 
Then add a new column that gives \% diameter relative to GY94 (the entry would just be 100\% for GY94).
So this gives 3 columns: model, tree diameter, \% of GY94 tree diameter}
} 
     \begin{tabular}{cccccccccc}
        \hline
         Model & no rate variation & {\shortstack{$\Gamma\omega$ rate\\ variation}} \\ \hline
         GY94 &53.05 & 72.28\\
         ExpCM(H1) & 57.32 & 81.31\\
         ExpCM(H3) & 57.77 & 78.98\\
         ExpCM(H1 + H3 avg) & 62.98 & 83.96 \\
      \end{tabular}
\end{supptable}

@@

\begin{suppfile}
\caption{
\label{suppfile:WSN_Perth_map}
List of alignable sites between H1 HA and H3 HA. 
This file provides a conversion between the numbering scheme we use in the paper (sequential numbering of just the alignable sites) to sequential numbering of the H1 HA reference sequence A/Wilson Smith/1933 and the H3 HA reference sequence A/Perth/2009. 
The sequential numbering for the HA strains starts with ``site 1" as the second codon (first non-methionine codon). 
\jdbcomment{The last sentence is so unclear that even I'm not sure what it means. If this isn't apparent from the file, explain it much better! Wouldn't it be simpler if you just used numbering that started with 1 at the first codon?
Also, add the actual file so we can see if it makes sense.}
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:H1_prefs}
Amino acid preferences measured by the deep mutational scanning of the H1 HA strain A/WSN/1933~\citep{doud2016accurate}. 
This file only contains measurements for the alignable sites between H1 and H3 HAs. 
Conversion from this numbering scheme to sequential numbering of A/WSN/1933 is in \ref{suppfile:WSN_Perth_map}. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:H3_prefs}
Amino acid preferences measured by the deep mutational scanning of the H3 HA strain A/Perth/2009~\citep{lee2018deep}. 
This file only contains measurements for the alignable sites between H1 and H3 HAs. 
Conversion from this numbering scheme to sequential numbering of A/Perth/2009 is in in \ref{suppfile:WSN_Perth_map}. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:avg_prefs}
Average of the amino acid preferences measured by the deep mutational scanning of the H1 HA strain A/Wilson Smith/1933~\citep{doud2016accurate} and the H3 HA strain A/Perth/2009~\citep{lee2018deep}. 
This file only contains measurements for the homologous sites between H1 HA and H3 HA. 
Conversion from this numbering scheme to sequential numbering in A/Wilson Smith/1933 or A/Perth/2009 is found in \ref{suppfile:WSN_Perth_map}. }
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:required_seqs}
This file contains the representative sequences of major human equine influenza lineages added to the alignment in \ref{suppfile:alignment_high}. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:alignment_high}
The HA sequences used for the full HA tree in \ref{fig:empirical_trees}. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
\jdbcomment{This legend is not very good. If the file just has the alignment, then why are we talking about RAxML in its legend?
Also, I think referring to \ref{fig:empirical_trees} to refer to the tree topology is not very clear, since that figure shows lots of stuff beyond just one tree.}
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:alignment_lowH1}
The HA sequences using for the ``low divergence from H1" tree in \ref{fig:compete}. 
All sequences have $\ge 59\%$ amino-acid identity to the H1 reference sequence A/Wilson Smith/1933. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:alignment_lowH3}
The HA sequences using for the ``low divergence from H3" tree in \ref{fig:compete}. 
All sequences have $\ge 59\%$ amino-acid identity to the H3 reference sequence A/Perth/2009. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:alignment_intermediateH1}
The HA sequences using for the ``intermediate divergence from H1" tree in \ref{fig:compete}. 
All sequences have $\ge 46\%$ amino-acid identity to the H1 reference sequence A/Wilson Smith/1933. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:alignment_intermediateH3}
The HA sequences using for the ``intermediate divergence from H3" tree in \ref{fig:compete}. 
All sequences have $\ge 46\%$ amino-acid identity to the H3 reference sequence A/Perth/2009. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_GY94}
The GY94 model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive}. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_GY94+Gr}
The GY94+$\Gamma\omega$ model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive}. 
}
\end{suppfile}

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_ExpCMdoud}
The ExpCM model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive} and the H1 HA amino-acid preference set in \ref{suppfile:H1_prefs}. 
}
\end{suppfile}

>>>>

\begin{figure}
\centerline{\includegraphics[width=0.99\textwidth]{figures/phylobayes.pdf}}
\caption{\label{fig:phylobayes}
\textbf{Models inferred from natural sequences have similar stationary states to models defined by experimental preferences and estimate similar branch lengths.}
(A) We fit an ExpCM(H1+H3 avg)+$\Gamma\omega$ and a pbMutSel to the full HA tree in \ref{fig:empirical_trees}. 
The pbMutSel amino-acid preferences are inferred from the natural HA sequences, while the ExpCM amino-acid preferences are experimentally measured and then rescaled by the stringency parameter in \ref{tab:empirical_data}. 
The pbMutSel preferences are more correlated with the re-scaled average of the H1 and H3 deep mutational scanning preferences than the individual re-scaled H1 and H3 deep mutational scanning preferences are to each other (Pearson's $r$: 0.74 versus 0.52). 
(B) The ExpCM(H1+H3 avg)+$\Gamma\omega$ and pbMutSel models estimated similar branch lengths when fit to the entire HA tree. 
\jdbcomment{You need to mention and give a bit of explanation for the coloring. What does it mean for a branch to be ``from'' H1 or H3? Would it be better to say branch ``leading to''? And does that mean leading to it with respect to the root of the tree as shown in \ref{fig:empirical_trees}?}
The \texttt{phydms} program implementing ExpCM's and the \texttt{PhyloBayes-MPI} program implementing pbMutSel models give branch lengths in different units, so to facilitate direct comparison between the models, we have normalized all branch lengths returned by each program by the length of the branches separating the earliest (A/South Carolina/1918) and latest (A/Solomon Islands/2006) seasonal human H1 sequences on the three. 
}
\end{figure}

The previous sections used ExpCM's, which are mutation-selection models that use site-specific amino-acid preferences that have been measured by experiments. 
However, there are other mathematically similar implementations of mutation-selection models that infer the amino-acid preferences directly from the natural sequence data. 
When these models are designed for use in phylogenetic inference, they are generally implemented in a Bayesian framework, which avoids the overfitting problems associated with trying to make maximum-likelihood estimates of the thousands of amino-acid preference parameters~\citep{lartillot2014overcoming}.
(Note that the maximum-likelihood implementations of \citet{tamuri2012estimating,tamuri2014penalized} are designed for estimating the amino-acid preferences, \emph{not} for phylogenetic inference.)
The model most comparable to our ExpCM's is the codon mutation-selection model implemented in \texttt{PhyloBayes-MPI}, which we will refer to as pbMutSel~\citep{rodrigue2014site}. 
In the pbMutSel model, the amino-acid preferences are modeled using Dirichlet processes rather than derived from experiments. 
However, like an ExpCM, a pbMutSel model still assumes a single set of time-homogeneous site-specific amino-acid preferences for the entire tree.

@@

We fit a pbMutSel model to the entire HA phylogenetic tree, and compared the results to those from analyzing the same tree with the best ExpCM, which is the ExpCM(H1+H3 avg)+$\Gamma\omega$ variant.
This is a direct apple-to-apples comparison, since the pbMutSel model also draws $\omega$ from a gamma-distribution~\citep{rodrigue2014site}.
First, we compared the amino-acid preferences inferred by the pbMutSel model to the preferences measured in the experiments.
\ref{fig:phylobayes}A shows that the preferences inferred by pbMutSel are quite similar to the (H1+ H3 avg) obtained by averaging the deep mutational scanning measurements for the H1 and H3 HAs. 
Notably, the amino-acid preferences from the pbMutSel model are more correlated with the (H1+ H3 avg) than the H1 and H3 measurements are with each other (\ref{fig:phylobayes}A).
This strong correlation indicates that the ExpCM(H1+H3 avg)+$\Gamma\omega$ is unlikely to be much different than a pbMutSel model that is parameterized only using the natural sequence data.

@@

We created four subalignments with ``low" and ``intermediate" divergence from either the H1 or the H3 deep mutational scanning reference sequence for the analysis in \ref{fig:compete}. 
The ``low divergence" alignments had $\ge 59\%$ amino-acid identity to the sequence used in the deep mutational scanning, and the ``intermediate divergence" alignments had $\ge 46\%$ identity from the reference sequence.
The resulting alignments are found in \ref{suppfile:alignment_lowH1}, \ref{suppfile:alignment_lowH3}, \ref{suppfile:alignment_intermediateH1}, and \ref{suppfile:alignment_intermediateH3}.  
\jdbcomment{Again, we should probably get rid of these four preceding files since we have too many. Presumably they can be re-generated using the description above of the identity cutoffs?}

@@

\subsection*{pbMutSel inference with \texttt{PhyloBayes-MPI}.}
For \ref{fig:phylobayes}, we fit a pbMutSel model to the full HA tree. 
We ran one chain for 5000 steps, saved every sample, and discarded the first 500 samples as a burnin. 
\jdbcomment{Presumably the program gives some measure of convergence? If so, describe what it was.}
\jdbcomment{You need to better describe what the tree represents and what the pbMutSel preferences represent. Are they the posterior mean?}

In order to make the branch lengths in \ref{fig:phylobayes} comparable between the pbMutSel tree returned by \texttt{PhyloBayes-MPI} and the other trees returned by \texttt{phydms}, we we normalized the branch lengths on the pbMutSel consensus tree and the ExpCM(H1+H3 avg)+$\Gamma\omega$ by dividing each branch by the length from A/South Carolina/1/1918 and A/Solomon Islands/3/2006. 
These two H1 sequences are early and late representatives of the longest known human influenza lineage, and are of sufficiently high identity that different ExpCM and GY94 substitution models all estimate nearly identical branch lengths separating them.

\subsection*{Software versions and computer code}
All code used for the analyses in this paper is available at \url{https://github.com/jbloomlab/divergence_timing_manuscript}. \skhcomment{make repo public}
The external computer programs that we used were
\begin{itemize} 
\item \texttt{phydms}~\citep{hilton2017phydms} version 2.2.2 (available at \url{github.com/jbloomlab/phydms}) to fit the ExpCM and GY94 models.
\item \texttt{pyvolve}~\citep{spielman2015pyvolve} version 0.8.7 (available at \url{https://github.com/sjspielman/pyvolve}) to simulate the sequences.
\item \texttt{PhyloBayes-MPI}~\citep{rodrigue2014site} version 1.8 (available at \url{https://github.com/bayesiancook/pbmpi}) to fit the pbMutSel model. 
\item \texttt{RAxML}~\citep{stamatakis2006raxml} version 8.2.11 (available at \url{https://github.com/stamatak/standard-RAxML} to infer tree topology.
\item We used \texttt{ggplot2}~\citep{wickham2016ggplot2}, \texttt{ggtree}~\citep{yu2017ggtree}, and \texttt{ggseqlogo}~\citep{wagih2017ggseqlogo} for visualization of the results.
\end{itemize}

@@

I would say keep the map of site numbers, the H1 and H3 preferences, and the alignment. 
I think all the other files can either be re-generated using the other ones if you provide adequate instructions in the methods (e.g., averaging the preferences or pruning the full alignment to sub-alignments), or simply have model parameters which could be put in a supplementary table.
}

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_doud}}
\caption{\label{suppfig:prefs_doud}
\textbf{H1 HA amino-acid preferences measured by deep mutational scanning.}
Each column represents a site in the HA protein, and the height of each letter is proportional to the preference for the amino acid measured by~\citet{doud2016accurate} and then re-scaled by the stringency parameter in \ref{tab:empirical_data}. 
The plot only shows sites that are alignable between the H1 and H3 HAs, and these alignable sites are numbered sequentially starting from 1.
The conversion between the numbering scheme in this figure and sequential numbering of the H1 HA reference sequence is in \ref{suppfile:WSN_Perth_map}. 
\jdbcomment{For this figure and the other logo plots, increase the number of rows by a few so things are a bit bigger.}
}
\end{suppfig}
\clearpage

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_lee}}
\caption{\label{suppfig:prefs_lee}
\textbf{H3 HA amino-acid preferences measured by deep mutational scanning.}
Similar to \ref{suppfig:prefs_doud} but shows the re-scaled preferences for the H3 HA as measured by \citet{lee2018deep}.
 }
\end{suppfig}
\clearpage

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_average}}
\caption{\label{suppfig:prefs_average}
\textbf{Average of H1 HA  and H3 HA amino-acid preferences measured by deep mutational scanning.}
Similar to \ref{suppfig:prefs_doud} but shows the re-scaled average of the preferences for the H1 and H3 HAs.
}
\end{suppfig}
\clearpage

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_mutSel}}
\caption{\label{suppfig:prefs_mutSel}
\textbf{Amino-acid preferences inferred by the pbMutSel model.}
Similar to \ref{suppfig:prefs_doud}, but shows the preferences inferred by fitting the pbMutSel model to the full HA tree.
}
\end{suppfig}
\clearpage

\begin{suppfig}[H]
\centerline{\includegraphics[width=0.50\textwidth]{figures/divergence_distances.pdf}}
\caption{\label{suppfig:subalignments}
\textbf{Overall divergence for the subtrees in \ref{fig:compete}.} 
\jdbcomment{This legend is indecipherable because rather than saying what you are showing you just refer to a lot of other figures / files.
Reduce the references to other figures / files to the minimum needed an explain what is actually being shown.}
We created two sub alignments from the alignment (\ref{suppfile:alignment_high}) used in \ref{fig:empirical_trees} for each deep mutational scanning reference HA. 
The ``low" alignments have $\ge59\%$ amino-acid identity to the reference sequence (\ref{suppfile:alignment_lowH1}, \ref{suppfile:alignment_lowH3}) and the ``intermediate" alignments have $\ge46\%$ amino-acid identity to the reference sequence (\ref{suppfile:alignment_intermediateH1}, \ref{suppfile:alignment_intermediateH3}).
}
\end{suppfig}
\clearpage

\begin{supptable}[t!]
\caption{\label{supptab:tree_lengths}
{\bf Total branch length for the trees in \ref{fig:empirical_trees}.}
The branch lengths are in average codon substitutions per site. 
\jdbcomment{Change to be tree diameter, which is distance between two most distant tip nodes.
Also, to make it easier to relate back to the percentage increases in the text, make 8 rows, one for each model and get rid of the ``rate variation'' column. 
Then add a new column that gives \% diameter relative to GY94 (the entry would just be 100\% for GY94).
So this gives 3 columns: model, tree diameter, \% of GY94 tree diameter}
} 
     \begin{tabular}{cccccccccc}
        \hline
         Model & no rate variation & {\shortstack{$\Gamma\omega$ rate\\ variation}} \\ \hline
         GY94 &53.05 & 72.28\\
         ExpCM(H1) & 57.32 & 81.31\\
         ExpCM(H3) & 57.77 & 78.98\\
         ExpCM(H1 + H3 avg) & 62.98 & 83.96 \\
      \end{tabular}
\end{supptable}
\clearpage

@@

\begin{suppfile}
\caption{
\label{suppfile:WSN_Perth_map}
List of alignable sites between H1 HA and H3 HA. 
This file provides a conversion between the numbering scheme we use in the paper (sequential numbering of just the alignable sites) to sequential numbering of the H1 HA reference sequence A/Wilson Smith/1933 and the H3 HA reference sequence A/Perth/2009. 
The sequential numbering for the HA strains starts with ``site 1" as the second codon (first non-methionine codon). 
\jdbcomment{The last sentence is so unclear that even I'm not sure what it means. If this isn't apparent from the file, explain it much better! Wouldn't it be simpler if you just used numbering that started with 1 at the first codon?
Also, add the actual file so we can see if it makes sense.}
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:H1_prefs}
Amino acid preferences measured by the deep mutational scanning of the H1 HA strain A/WSN/1933~\citep{doud2016accurate}. 
This file only contains measurements for the alignable sites between H1 and H3 HAs. 
Conversion from this numbering scheme to sequential numbering of A/WSN/1933 is in \ref{suppfile:WSN_Perth_map}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:H3_prefs}
Amino acid preferences measured by the deep mutational scanning of the H3 HA strain A/Perth/2009~\citep{lee2018deep}. 
This file only contains measurements for the alignable sites between H1 and H3 HAs. 
Conversion from this numbering scheme to sequential numbering of A/Perth/2009 is in in \ref{suppfile:WSN_Perth_map}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:avg_prefs}
Average of the amino acid preferences measured by the deep mutational scanning of the H1 HA strain A/Wilson Smith/1933~\citep{doud2016accurate} and the H3 HA strain A/Perth/2009~\citep{lee2018deep}. 
This file only contains measurements for the homologous sites between H1 HA and H3 HA. 
Conversion from this numbering scheme to sequential numbering in A/Wilson Smith/1933 or A/Perth/2009 is found in \ref{suppfile:WSN_Perth_map}. }
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:required_seqs}
This file contains the representative sequences of major human equine influenza lineages added to the alignment in \ref{suppfile:alignment_high}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_high}
The HA sequences used for the full HA tree in \ref{fig:empirical_trees}. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
\jdbcomment{This legend is not very good. If the file just has the alignment, then why are we talking about RAxML in its legend?
Also, I think referring to \ref{fig:empirical_trees} to refer to the tree topology is not very clear, since that figure shows lots of stuff beyond just one tree.}
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_lowH1}
The HA sequences using for the ``low divergence from H1" tree in \ref{fig:compete}. 
All sequences have $\ge 59\%$ amino-acid identity to the H1 reference sequence A/Wilson Smith/1933. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_lowH3}
The HA sequences using for the ``low divergence from H3" tree in \ref{fig:compete}. 
All sequences have $\ge 59\%$ amino-acid identity to the H3 reference sequence A/Perth/2009. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_intermediateH1}
The HA sequences using for the ``intermediate divergence from H1" tree in \ref{fig:compete}. 
All sequences have $\ge 46\%$ amino-acid identity to the H1 reference sequence A/Wilson Smith/1933. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:alignment_intermediateH3}
The HA sequences using for the ``intermediate divergence from H3" tree in \ref{fig:compete}. 
All sequences have $\ge 46\%$ amino-acid identity to the H3 reference sequence A/Perth/2009. 
We inferred the topology using \texttt{RAxML} and the GTRCAT model. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_GY94}
The GY94 model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_GY94+Gr}
The GY94+$\Gamma\omega$ model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive}. 
}
\end{suppfile}
\clearpage

\begin{suppfile}
\caption{
\label{suppfile:hybrid_lowH1_ExpCMdoud}
The ExpCM model parameters fit to the ``low divergence from H1" tree (\ref{fig:compete}, \ref{suppfile:alignment_lowH1}). 
The model was fit using \texttt{phydms\_comprehensive} and the H1 HA amino-acid preference set in \ref{suppfile:H1_prefs}. 
}
\end{suppfile}
\clearpage
